---
title: "Day 7 Homework"
output: html_document
---

1. Work on making progress on your Final Project.

# Tonight I am reading about the fda package in R and the techniques I am going to try to use

2. Document some road blocks you are running into.

# I have spent most of my time cleaning and settign up my data. Two important issues I faced was merging all 11 datasets into one longitudinal file, and making that file in long format. I was able to merge successfully (see code below) but after numerous hours, I have still be unable to successfully melt the longitudinal data file. I tried just the numeric data, but it is just showing the last follow-up period, despite saying there are 8,855,160 observations (see code below). 

#merge and follow-up indicator variables
```{r}
pathbase0 <- da29961.0001
pathbase0$time <- 0
pathbase0$followup <- "baseline"


pathbase06 <- da29961.0002
pathbase06$time <- 6
pathbase06$followup <- "6-month"

pathbase12 <- da29961.0003
pathbase12$time <- 12
pathbase12$followup <- "12-month"

pathbase18 <- da29961.0004
pathbase18$time <- 18
pathbase18$followup <- "18-month"

pathbase24 <- da29961.0005
pathbase24$time <- 24
pathbase24$followup <- "24-month"

pathbase30 <- da29961.0006
pathbase30$time <- 30
pathbase30$followup <- "30-month"

pathbase36 <- da29961.0007
pathbase36$time <- 36
pathbase36$followup <- "36-month"

pathbase48 <- da29961.0008
pathbase48$time <- 48
pathbase48$followup <- "48-month"

pathbase60 <- da29961.0009
pathbase60$time <- 60
pathbase60$followup <- "60-month"

pathbase72 <- da29961.0010
pathbase72$time <- 72
pathbase72$followup <- "72-month"

pathbase84 <- da29961.0011
pathbase84$time <- 84
pathbase84$followup <- "84-month"

# my first try (didn't work)

pathbaseall <- merge(pathbase0, pathbase06, by = "CASEID", all = TRUE)
pathbaseall1 <- merge(pathbaseall, pathbase12, by = "CASEID", all = TRUE)
pathbaseall2 <- merge(pathbaseall1, pathbase18, by = "CASEID", all = TRUE) 
pathbaseall3 <- merge(pathbaseall2, pathbase24, by = "CASEID", all = TRUE)
pathbaseall4 <- merge(pathbaseall3, pathbase30, by = "CASEID", all = TRUE)
pathbaseall5 <- merge(pathbaseall4, pathbase36, by = "CASEID", all = TRUE)
pathbaseall6 <- merge(pathbaseall5, pathbase48, by = "CASEID", all = TRUE)
pathbaseall7 <- merge(pathbaseall6, pathbase60, by = "CASEID", all = TRUE)
pathbaseall8 <- merge(pathbaseall7, pathbase72, by = "CASEID", all = TRUE)
pathbase_all <- merge(pathbaseall8, pathbase84, by = "CASEID", all = TRUE)

# my second try (also didn't work)

run.seq <- function(x) as.numeric(ave(paste(x), x, FUN = seq_along))

L <- list(pathbase0, pathbase06, pathbase12, pathbase18, pathbase24, pathbase30, pathbase36, pathbase48, pathbase60, pathbase72, pathbase84)
merge_attempt<-ldply(L, id.vars = c("CASEID", "time"))
L2 <- lapply(L, function(x) cbind(x, run.seq = run.seq(c(x$time, x$CASID))))
out <- Reduce(function(...) merge(..., all = TRUE), L2)[-2]

```

# Attempt at melting the longitudinal data file (only numeric variables). Unsurprisingly, it also did not work 
```{r}


numeric.cols <- rep(NA, ncol(out))
for(colnum in 1:ncol(out)) {
  numeric.cols[colnum] <- is.numeric(out[, colnum])
}
numeric.cols[1:20]

pathbase_numeric <- out[, numeric.cols]

pathbase_all.long <- melt(pathbase_numeric, id.vars = c("CASEID", "time"))
colnames(pathbase_all)
head(pathbase_all_num.long)
tail(pathbase_all_num.long)
pathbase_all$time
```

3. Email them to me before 7am tomorrow with your interests for special topics day. 